---
title: "capstone_xz_ml"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
install.packages('dlookr')
library(dlookr)
install.packages('RCurl')
install.packages('MASS')
install.packages('leaps')
library(RCurl) # getURL 
library(MASS) # stepwise regression
library(leaps) # all subsets regression
```

# read *norm file, 
```{r}
df_norm <- read.csv(file ='c:/work/source/cell5m_17_na_rm_norm_binnedTarget.csv',stringsAsFactors = TRUE)
df_norm$vp_n_bin <- factor(df_norm$vp_n_bin, order = TRUE, levels = c('[0,0.0211]', '(0.0211,0.0376]', '(0.0376,0.0607]', '(0.0607,1]'))
str(df_norm)
summary(df_norm)
```


## subset data

```{r}
data_train <- sample(nrow(df_norm), floor(nrow(df_norm)*0.7))
train <- df_norm[data_train,]
test <- df_norm[-data_train,]
```

ordinal logistic regression model to predict 'vp' variable. 
```{r}
library(MASS)
model_polr_norm <- polr(vp_n_bin~.-X-CELL5M-x-y, data=train, Hess = TRUE)
summary(model_polr_norm)
prediction_norm <- predict(model_polr_norm, interval="prediction", newdata =test)
```
Calculate the significance of coefficients and intercepts
```{r}
summary_table <- coef(summary(model_polr_norm))
pval <- pnorm(abs(summary_table[, "t value"]),lower.tail = FALSE)* 2
summary_table <- cbind(summary_table, "p value" = round(pval,3))
summary_table

```
#Forward selection algorithm. start with 'null', to come up with a selection of independent variables between 'null' and 'full'. 
```{r echo=TRUE}
library(MASS)
full_norm <- polr(vp_n_bin~.-X-CELL5M-x-y, data=df_norm)
null_norm <- polr(vp_n_bin~1,data=df_norm)
stepF_norm <- stepAIC(null_norm, scope=list(lower=null_norm, upper=full_norm), direction= "forward", trace=TRUE)
summary(stepF_norm)
```

#'backward' elimination, which will start with 'full'.
```{r echo=TRUE}
full_norm <- polr(vp_n_bin~.-X-CELL5M-x-y, data=df_norm)
stepB_norm <- stepAIC(full_norm, direction= "backward", trace=TRUE)
summary(stepB_norm)
```

#test the best combination of the attributes.
```{r}
library(leaps)
subsets<-regsubsets(vp_n_bin~.-X-CELL5M-x-y, data=df_norm, nbest=1,)
sub.sum <- summary(subsets)
as.data.frame(sub.sum$outmat)


```
In the output * denotes the included variables.

The best combination of attributes is: 'pre_mean', 'TT_20K', and some factors of 'AEZ8_CLAS' and 'FS_2012_TX'
however, factor attributes 'AEZ8_CLAS' and 'FS_2012_TX' accounts for 20 out of 30 attributes, it seems to overwhelm the other attributes.  In this context, it is important to note that 'pre_mean' and 'TT_20K' still stand out as very significant attributes.









