---
title: "capstone"
output: word_document
---

#install R packages
# define RTOOL path
```{r}
#install.packages('RTool')
library(RTool)
#writeLINES('PATH="${RTooLS40_HOME}\\usr\\bin;${PATH}"', con = "~/Reviron")
#Sys.which("make")
## "C:\\rtools40\\usr\\bin\\make.exe"

install.packages('plyr')
library(plyr)
install.packages('dplyr')
library(dplyr)
install.packages('ggplot2')
library(ggplot2)
install.packages('sqldf')
library(sqldf)
install.packages('dlookr')
library(dlookr)
install.packages("Hmisc")
library(Hmisc)
```
#read files
```{r}
#data <- read.csv(file ='c:/work/CELL5M_data.csv')
#str(data)
Vop <- read.csv(file ='c:/work/source/Agriculture_Crops_vopagg.csv')
Zone <- read.csv(file ='c:/work/source/Agroecology_Agro-EcologicalZones.csv')
Climate <- read.csv(file ='c:/work/source/Agroecology_Climate.csv')
Elev <- read.csv(file ='c:/work/source/Agroecology_Elevation.csv')
FmSys <- read.csv(file ='c:/work/source/Agroecology_FarmingSystems.csv')
Soil <- read.csv(file ='c:/work/source/Agroecology_SoilResources.csv')
Income <- read.csv(file ='c:/work/source/Demographics_Income.csv')
Population <- read.csv(file ='c:/work/source/Demographics_Population.csv')  
Market <- read.csv(file ='c:/work/source/markets_foxp.csv')

```

#subset 
```{r}
library(dplyr)
Vop_sub <- Vop %>%
  select(cell5m, x, y, vp_cr_ar_r)
Zone_sub <- Zone %>%
  select(CELL5M, AEZ8_CLAS, AEZ16_CLAS)
Climate_sub <- Climate %>%
  select(CELL5M, pre_mean, LGP_AVG)
Elev_sub <- Elev %>%
  select(CELL5M, ELEVATION, SLOPE)
FmSys_sub <- FmSys %>%
  select(CELL5M, FS_2012_TX)
Soil_sub <- Soil %>%
  select(CELL5M, soc_d5, soc_d15, soc_d30, soc_d60, soc_d100, soc_d200)
Population_sub <- Population %>%
  select(CELL5M, PN05_TOT, PN05_RUR)
Income_sub <- Income %>%
  select(CELL5M, RPCX08_11)
Market_sub <- Market %>%
  select(CELL5M, TT_PORT, TT_20K)


```

#change Vop_sub primary key names
```{r}
Vop_sub <- rename(Vop_sub, CELL5M = cell5m)
```

#left join to Vop_sub
```{r}
vz <- left_join(x = Vop_sub, y = Zone_sub, by = "CELL5M")
vzc <-  left_join(x = vz, y = Climate_sub, by = "CELL5M")
vzce <- left_join(x = vzc, y = Elev_sub, by = "CELL5M")
vzcef <- left_join(x = vzce, y = FmSys_sub, by = "CELL5M" )
vzcefs <- left_join(x = vzcef, y = Soil_sub, by = "CELL5M")
vzcefsp <- left_join(x = vzcefs, y = Population_sub, by = "CELL5M")
vzcefspi <- left_join(x = vzcefsp, y = Income_sub, by = "CELL5M")
vzcefspim <- left_join(x = vzcefspi, y = Market_sub, by = "CELL5M")
str(vzcefspim)
write.csv(vzcefspim, 'c:/work/source/cell5m_22.csv')
data22 <- vzcefspim
```

#assign character attributes to factor, check data types
```{r}
data22$AEZ16_CLAS = factor(data22$AEZ16_CLAS)
data22$AEZ8_CLAS = factor(data22$AEZ8_CLAS)
data22$FS_2012_TX = factor(data22$FS_2012_TX)
sapply(data22, class)
```

#subset rows by filtering dependent variable = 0

```{r}
data22_agro <- data22 %>% filter(vp_cr_ar_r > 0)
str(data22_agro)
sum(is.na(data22_agro$AEZ8_CLAS) == TRUE)
sum(is.na(data22_agro$AEZ16_CLAS) == TRUE)
sum(is.na(data22_agro$pre_mean) == TRUE)
sum(is.na(data22_agro$LGP_AVG) == TRUE)
sum(is.na(data22_agro$ELEVATION) == TRUE)
sum(is.na(data22_agro$SLOPE) == TRUE)
sum(is.na(data22_agro$FS_2012_TX) == TRUE)
sum(is.na(data22_agro$soc_d5) == TRUE)
sum(is.na(data22_agro$soc_d15) == TRUE)
sum(is.na(data22_agro$soc_d30) == TRUE)
sum(is.na(data22_agro$soc_d60) == TRUE)
sum(is.na(data22_agro$soc_d100) == TRUE)
sum(is.na(data22_agro$soc_d200) == TRUE)
sum(is.na(data22_agro$PN05_TOT) == TRUE)
sum(is.na(data22_agro$PN05_RUR) == TRUE)
sum(is.na(data22_agro$UPCX08_11) == TRUE)
sum(is.na(data22_agro$RPCX08_11) == TRUE)
sum(is.na(data22_agro$TT_PORT) == TRUE)
sum(is.na(data22_agro$TT_20K) == TRUE)
```

#remove income RPCX08_11
```{r}
data21_agro <- data22_agro[, -20]

```

#remove rows with at least one missing values
```{r}
data21_agro_na <- rowSums(is.na(data21_agro))
data21_agro_1na <- data21_agro_na >= 1
data21_agro_na_rm <- data21_agro[!data21_agro_1na,]
str(data21_agro_na_rm)
write.csv(data21_agro_na_rm, 'c:/work/source/cell5m_21_na_rm.csv')
```

#histogram for dependent variable

```{r}
hist(data21_agro_na_rm$vp_cr_ar_r, main = "Histogram: Crop Value Per Ha (Dependent Variable)")

```
#univariant box plots on log(10) scale
```{r}
library(ggplot2)
ggplot(data21_agro_na_rm, aes(y=vp_cr_ar_r), main = "Boxplot: Crop Value Per Ha (Dependent Variable)")+geom_boxplot()+scale_y_log10()
ggplot(data21_agro_na_rm, aes(y=TT_PORT))+geom_boxplot()
ggsave("boxplot_vp_cr_ar_r.png", width=5, height = 5)

```

#bivariant box plots 
```{r}
boxplot(vp_cr_ar_r~AEZ16_CLAS, data=data21_agro_na_rm, main = "Box Plots: Crop Values of Agroecological Zones")
boxplot(vp_cr_ar_r~FS_2012_TX, data=data21_agro_na_rm, main = "Box Plots: Crop Values of Farming Systems")

```

#bivariant: scatter plots
```{r}
ggplot(data21_agro_na_rm, aes(SLOPE, vp_cr_ar_r)) + geom_point() + scale_y_log10() + scale_x_log10()
ggplot(data21_agro_na_rm, aes(pre_mean, vp_cr_ar_r)) + geom_point() + scale_y_log10() 
ggplot(data21_agro_na_rm, aes(LGP_AVG, pre_mean)) + geom_point() + scale_y_log10() + scale_x_log10()
ggplot(data21_agro_na_rm, aes(LGP_AVG, ELEVATION)) + geom_point() + scale_y_log10() + scale_x_log10()
ggplot(data21_agro_na_rm, aes(pre_mean, ELEVATION)) + geom_point() + scale_y_log10() + scale_x_log10()
ggplot(data21_agro_na_rm, aes(soc_d100, soc_d200)) + geom_point() + scale_y_log10() + scale_x_log10()
```

#scatter plots
```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor) 
{
    usr <- par("usr"); on.exit(par(usr)) 
    par(usr = c(0, 1, 0, 1)) 
    r <- abs(cor(x, y)) 
    txt <- format(c(r, 0.123456789), digits=digits)[1] 
    txt <- paste(prefix, txt, sep="") 
    if(missing(cex.cor)) cex <- 0.8/strwidth(txt) 
 
    test <- cor.test(x,y) 
    # borrowed from printCoefmat
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE, 
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " ")) 
 
    text(0.5, 0.5, txt, cex = cex * r) 
    text(.8, .8, Signif, cex=cex, col=2)
}
selFea2 <- data21_agro_na_rm[,	c('vp_cr_ar_r', 'soc_d5', 'soc_d30','soc_d60')]
selFea3 <- data21_agro_na_rm[,	c('vp_cr_ar_r', 'soc_d60', 'soc_d100','soc_d200')]
selFea4 <- data21_agro_na_rm[,	c('vp_cr_ar_r', 'pre_mean', 'LGP_AVG', 'ELEVATION')]
selFea5 <- data21_agro_na_rm[,	c('vp_cr_ar_r', 'PN05_TOT', 'PN05_RUR', 'TT_PORT', 'TT_20K')]
selFea6 <- data21_agro_na_rm[,	c('vp_cr_ar_r', 'AEZ8_CLAS', 'FS_2012_TX', 'TT_PORT', 'TT_20K')]
pairs(selFea2, lower.panel=panel.smooth, upper.panel=panel.cor)
pairs(selFea3, lower.panel=panel.smooth, upper.panel=panel.cor)
pairs(selFea4, lower.panel=panel.smooth, upper.panel=panel.cor)
pairs(selFea5, lower.panel=panel.smooth, upper.panel=panel.cor)
pairs(selFea6, lower.panel=panel.smooth, upper.panel=panel.cor)


```

#scatter plot on farming systems
```{r}
qplot(x = AEZ8_CLAS, y = AEZ16_CLAS, data= data21_agro_na_rm) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust =1))

```
#univariant: bar plots
```{r}
ggplot(data21_agro_na_rm, aes(x=factor(AEZ16_CLAS))) + geom_bar(stat="count", width=0.7, fill="steelblue")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust =1))
ggplot(data21_agro_na_rm, aes(x=factor(AEZ8_CLAS))) + geom_bar(stat="count", width=0.7, fill="steelblue")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust =1))
ggplot(data21_agro_na_rm, aes(x=factor(FS_2012_TX))) + geom_bar(stat="count", width=0.7, fill="steelblue")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust =1))

```
#remove highly correlated / redundant features: soc_d15, soc_d30, soc_d100
#remove redundant attribute:AEZ16_CLAS

```{r}
sub1 <- data21_agro_na_rm[, -13]
sub2 <- sub1[, -13]
sub3 <- sub2[, -14]
sub4 <- sub3[, -6]
str(sub4)
summary(sub4)
```
#save to file for further analysis
```{r}
write.csv(sub4, 'c:/work/source/cell5m_17_na_rm.csv')

```

#normalization: min-max function
```{r}
normalize <- function(x) {
               return ((x - min(x)) / (max(x) - min(x))) }

sub4a_n <- as.data.frame(lapply(sub4[4], normalize))
sub4b_n <- as.data.frame(lapply(sub4[6:9], normalize))
sub4c_n <- as.data.frame(lapply(sub4[11:17], normalize))
sub4_n <- cbind(sub4[1:3], sub4a_n, sub4b_n, sub4c_n, sub4[5], sub4[10])
str(sub4_n)
summary(sub4_n)
hist(sub4_n$vp_cr_ar_r)
var(sub4_n$vp_cr_ar_r)
hist(sub4_n$pre_mean)
var(sub4_n$pre_mean)
hist(sub4_n$LGP_AVG)
var(sub4_n$LGP_AVG)
hist(sub4_n$ELEVATION)
var(sub4_n$ELEVATION)
hist(sub4_n$SLOPE)
var(sub4_n$SLOPE)
hist(sub4_n$soc_d5)
var(sub4_n$soc_d5)
hist(sub4_n$soc_d60)
var(sub4_n$soc_d60)
hist(sub4_n$soc_d200)
var(sub4_n$soc_d200)
hist(sub4_n$PN05_TOT)
var(sub4_n$PN05_TOT)
hist(sub4_n$PN05_RUR)
var(sub4_n$PN05_RUR)
hist(sub4_n$TT_PORT)
var(sub4_n$TT_PORT)
hist(sub4_n$TT_20K)
var(sub4_n$TT_20K)
```

#save to file, *norm.csv
```{r}
write.csv(sub4_n, 'c:/work/source/cell5m_17_na_rm_norm.csv')

```

#log+1 transformation
#Note: after normalization to 0 - 1, log+1 no longer make sense
```{r}
vp_lg1 <- transform(sub4_n$vp_cr_ar_r, method=c("log+1"))
hist(sub4_n$vp_cr_ar_r)
hist(vp_lg1)
pre_lg1 <- transform(sub4_n$pre_mean, method=c("log+1"))
hist(pre_lg1)
elev_lg1 <- transform(sub4_n$ELEVATION, method=c("log+1"))
hist(elev_lg1)
#produced 1 NA, remove NA
elev_lg1_na <- is.na(elev_lg1)
#replace 1 na to 0 
elev_lg1[which(is.na(elev_lg1))] = 0
#check
elev_lg1_na <- is.na(elev_lg1)
sum(elev_lg1_na)

slop_lg1 <- transform(sub4_n$SLOPE, method=c("log+1"))
hist(slop_lg1)
soc5_lg1 <- transform(sub4_n$soc_d5, method=c("log+1"))
hist(soc5_lg1)
soc60_lg1 <- transform(sub4_n$soc_d60, method=c("log+1"))
hist(soc60_lg1)
soc200_lg1 <- transform(sub4_n$soc_d200, method=c("log+1"))
hist(soc200_lg1)
PNT_lg1 <- transform(sub4_n$PN05_TOT, method=c("log+1"))
hist(PNT_lg1)
PNR_lg1 <- transform(sub4_n$PN05_RUR, method=c("log+1"))
hist(PNT_lg1)
TPORT_lg1 <- transform(sub4_n$TT_PORT, method=c("log+1"))
hist(TPORT_lg1)
T20k_lg1 <- transform(sub4_n$TT_20K, method=c("log+1"))
hist(T20k_lg1)
hist(sub4$TT_20K)


PNR_lg1 <- transform(sub4$PN05_RUR, method=c("log+1"))
hist(PNT_lg1)
var(PNT_lg1)
```
#standardization or zscore normalization, output file *stand.csv
```{r}
standardize <- function(x) {
               return ((x - mean(x)) / sd(x))}

sub4a_s <- as.data.frame(lapply(sub4[4], standardize))
sub4b_s <- as.data.frame(lapply(sub4[6:9], standardize))
sub4c_s <- as.data.frame(lapply(sub4[11:17], standardize))
sub4_s <- cbind(sub4[1:3], sub4a_s, sub4b_s, sub4c_s, sub4[5], sub4[10])
str(sub4_s)
summary(sub4_s)
write.csv(sub4_s, 'c:/work/source/cell5m_17_na_rm_stand.csv')
```

#sqrt transformation
```{r}
elev_sqrt <- transform(sub4_n$ELEVATION, method=c("sqrt"))
hist(elev_sqrt)
#na produced, check
elev_sqrt_na <- is.na(elev_sqrt)
sum(elev_sqrt_na)
#replace 1 na with 0
elev_sqrt[which(is.na(elev_sqrt))] = 0
#check
elev_sqrt_na <- is.na(elev_sqrt)
sum(elev_sqrt_na)
slop_sqrt <- transform(sub4_n$SLOPE, method=c("sqrt"))
hist(slop_sqrt)

```

#check population variance and distribution, after normalization/standardization/lg and sqrt transformed
```{r}
var(sub4_n$SLOPE)
var(sub4_s$SLOPE)
var(slop_lg1)
var(slop_sqrt)
hist(sub4_n$SLOPE)
hist(sub4_s$SLOPE)
hist(slop_lq1)
hist(slop_sqrt)
var(sub4_n$ELEVATION)
var(sub4_s$ELEVATION)
var(elev_lg1)
var(elev_sqrt)
hist(sub4_n$ELEVATION)
hist(sub4_s$ELEVATION)
hist(elev_lq1)
hist(elev_sqrt)
var(sub4_n$PN05_TOT)
var(sub4_s$PN05_TOT)
var(PNT_lg1)
hist(sub4_n$PN05_TOT)
hist(sub4_s$PN05_TOT)
hist(PNT_lg1)
var(sub4_n$PN05_RUR)
var(sub4_s$PN05_RUR)
var(PNR_lg1)
hist(sub4_n$PN05_RUR)
hist(sub4_s$PN05_RUR)
hist(PNR_lg1)
var(sub4_n$vp_cr_ar_r)
var(sub4_s$vp_cr_ar_r)
var(vp_lg1)
hist(sub4_n$vp_cr_ar_r)
hist(sub4_s$vp_cr_ar_r)
hist(vp_lg1)
var(sub4_n$pre_mean)
var(sub4_s$pre_mean)
hist(sub4_n$pre_mean)
hist(sub4_s$pre_mean)
var(sub4_n$LGP_AVG)
var(sub4_s$LGP_AVG)
hist(sub4_n$LGP_AVG)
hist(sub4_s$LGP_AVG)
var(sub4_n$soc_d5)
var(sub4_s$soc_d5)
var(soc5_lg1)
hist(sub4_n$soc_d5)
hist(sub4_s$soc_d5)
hist(soc5_lg1)
var(sub4_n$soc_d60)
var(sub4_s$soc_d60)
var(soc60_lg1)
hist(sub4_n$soc_d60)
hist(sub4_s$soc_d60)
hist(soc60_lg1)
var(sub4_n$soc_d200)
var(sub4_s$soc_d200)
var(soc200_lg1)
hist(sub4_n$soc_d200)
hist(sub4_s$soc_d200)
hist(soc200_lg1)
var(sub4_n$TT_PORT)
var(sub4_s$TT_PORT)
var(TPORT_lg1)
hist(sub4_n$TT_PORT)
hist(sub4_s$TT_PORT)
hist(TPORT_lg1)
var(sub4_n$TT_20K)
var(sub4_s$TT_20K)
var(T20k_lg1)
hist(sub4_n$TT_20K)
hist(sub4_s$TT_20K)
hist(T20K_lg1)


```

#create transformed df1 (without factor attributes), df2 (all 17 attributes). output df1 and df2 *tran.csv
```{r}

AEZ8_CLAS <- sub4_n$AEZ8_CLAS
pre_mean <- sub4_n$pre_mean
LGP_AVG <- sub4_n$LGP_AVG
FS_2012_TX <- sub4_n$FS_2012_TX
df1 <- cbind(sub4_n[, 1:3], vp_lg1, pre_mean, LGP_AVG, elev_sqrt, slop_sqrt, soc5_lg1, soc60_lg1, soc200_lg1, PNT_lg1, PNR_lg1, TPORT_lg1, T20k_lg1)
str(df1)
summary(df1)
write.csv(df1, 'c:/work/source/cell5m_15_na_rm_tran.csv')

df2 <- cbind(sub4_n[, 1:3], vp_lg1, AEZ8_CLAS, pre_mean, LGP_AVG, elev_sqrt, slop_sqrt, FS_2012_TX, soc5_lg1, soc60_lg1, soc200_lg1, PNT_lg1, PNR_lg1, TPORT_lg1, T20k_lg1)
str(df2)
summary(df2)
write.csv(df2, 'c:/work/source/cell5m_17_na_rm_tran.csv')
```

#change dependent attribute to factors using quantile binning from package dlookr: original, normalized (n), log+1 & sqrt  transformed (lg1) and standardized (s)
```{r}
library(dlookr)
vp_bin <- binning (
  sub4$vp_cr_ar_r,
  4, 
  type = c("quantile"),
  ordered = TRUE,
  labels = c("Q1", "Q2", "Q3", "Q4"),
  approxy.lab = TRUE
)
summary(vp_bin)

vp_n_bin <- binning (
  sub4_n$vp_cr_ar_r,
  4, 
  type = c("quantile"),
  ordered = TRUE,
  labels = NULL,
  approxy.lab = TRUE
)
summary(vp_n_bin)


vp_s_bin <- binning (
  sub4_s$vp_cr_ar_r,
  4, 
  type = c("quantile"),
  ordered = TRUE,
  labels = NULL,
  approxy.lab = TRUE
)
summary(vp_s_bin)

vp_lg1_bin <- binning (
  df2$vp_lg1,
  4, 
  type = c("quantile"),
  ordered = TRUE,
  labels = NULL,
  approxy.lab = TRUE
)
summary(vp_lg1_bin)
```

create new dataframes with binned dependent attributes
```{r}
df3 <- cbind(sub4_n[, 1:3], vp_lg1_bin, pre_mean, LGP_AVG, elev_sqrt, slop_sqrt, soc5_lg1, soc60_lg1, soc200_lg1, PNT_lg1, PNR_lg1, TPORT_lg1, T20k_lg1)
str(df3)
summary(df3)
write.csv(df3, 'c:/work/source/cell5m_15_na_rm_tran_binnedTarget.csv')

df4 <- cbind(sub4_n[, 1:3], vp_lg1_bin,  pre_mean, LGP_AVG, elev_sqrt, slop_sqrt, soc5_lg1, soc60_lg1, soc200_lg1, PNT_lg1, PNR_lg1, TPORT_lg1, T20k_lg1, AEZ8_CLAS,FS_2012_TX)
str(df4)
summary(df4)
write.csv(df4, 'c:/work/source/cell5m_17_na_rm_tran_binnedTarget.csv')

df5 <- cbind(sub4_n[, 1:3], vp_n_bin, sub4b_n, sub4c_n, sub4[5], sub4[10])
write.csv(df5, 'c:/work/source/cell5m_17_na_rm_norm_binnedTarget.csv')

df6 <- cbind(sub4_n[, 1:3], vp_n_bin, sub4b_n, sub4c_n)
write.csv(df6, 'c:/work/source/cell5m_15_na_rm_norm_binnedTarget.csv')

df7 <- cbind(sub4_s[, 1:3], vp_s_bin, sub4b_s, sub4c_s, sub4[5], sub4[10])
write.csv(df7, 'c:/work/source/cell5m_17_na_rm_stand_binnedTarget.csv')
df8 <- cbind(sub4_s[, 1:3], vp_s_bin, sub4b_s, sub4c_s)
write.csv(df8, 'c:/work/source/cell5m_15_na_rm_stand_binnedTarget.csv')

str(sub4)
df8 <- cbind(sub4[, 1:3], vp_bin, sub4[, 5:17])
write.csv(df8, 'c:/work/source/cell5m_17_na_rm_binnedTarget.csv')

```
#binning independent attributes 
(this is not used )
```{r}
pre_mean_bin <- binning (
  df2$pre_mean,
  4, 
  type = c("quantile"),
  ordered = TRUE,
  labels = NULL,
  approxy.lab = TRUE
)
summary(pre_mean_bin)

LGP_AVG_bin <- binning (
  df2$LGP_AVG,
  4, 
  type = c("quantile"),
  ordered = TRUE,
  labels = NULL,
  approxy.lab = TRUE
)
summary(LGP_AVG_bin)

elev_sqrt_bin <- binning (
  df2$elev_sqrt,
  4, 
  type = c("quantile"),
  ordered = TRUE,
  labels = NULL,
  approxy.lab = TRUE
)
summary(elev_sqrt_bin)

slop_sqrt_bin <- binning (
  df2$slop_sqrt,
  4, 
  type = c("quantile"),
  ordered = TRUE,
  labels = NULL,
  approxy.lab = TRUE
)
summary(slop_sqrt_bin)

soc5_lg1_bin <- binning (
  df2$soc5_lg1,
  4, 
  type = c("quantile"),
  ordered = TRUE,
  labels = NULL,
  approxy.lab = TRUE
)
summary(soc5_lg1_bin)

soc60_lg1_bin <- binning (
  df2$soc60_lg1,
  4, 
  type = c("quantile"),
  ordered = TRUE,
  labels = NULL,
  approxy.lab = TRUE
)
summary(soc60_lg1_bin)

soc200_lg1_bin <- binning (
  df2$soc200_lg1,
  4, 
  type = c("quantile"),
  ordered = TRUE,
  labels = NULL,
  approxy.lab = TRUE
)
summary(soc200_lg1_bin)

```



