
---
title: "K-Nearest Neighbours, normalized data set"
output: word_document
---


```{r setup, include=FALSE}
install.packages("class")
install.packages("gmodels")
library("class")
library("gmodels")
```

Normalized data set will be used.  Factor attributes are excluded.
```{r}
df_norm_bin <- read.csv(file ='c:/work/source/cell5m_15_na_rm_norm_binnedTarget.csv')
str(df_norm_bin)
```

subset data into training and test set.
```{r}
train_index <- sample(1:nrow(df_norm_bin), 0.7 * nrow(df_norm_bin))
train.set <- df_norm_bin[train_index,]
test.set  <- df_norm_bin[-train_index,]
```

remove target and spatial index columns from the training and test datasets.
```{r}
train.set_n1 <- train.set[-1]
train.set_n2 <- train.set_n1[-2]
train.set_n3 <- train.set_n2[-3]
train.set_n4 <- train.set_n3[-4]
train.set_new <- train.set_n4[-5]
test.set_n1 <- test.set[-1]
test.set_n2 <- test.set_n1[-2]
test.set_n3 <- test.set_n2[-3]
test.set_n4 <- test.set_n3[-4]
test.set_new <- test.set_n4[-5]
```

store the labels from our training and test datasets.
```{r}
df_train_labels <- train.set$vp_n_bin 
df_test_labels <- test.set$vp_n_bin
```

For k=4, make predication
```{r}
df_knn_prediction <- knn(train = train.set_new, test = test.set_new, cl= df_train_labels, k = 4) 
```

Let's see the confusion matrix.
```{r}
CrossTable(x=df_test_labels, y=df_knn_prediction, prop.chisq=FALSE)
```
to do: calculate the following for the model
accuracy
error rate
precision
sensitivity (recall)
specificity

TP = 25986
FN = 11467
FP = 11467
TN = ? is it possible to have total TN?



